House price data in Brunei

### 2024-12-14 analysis.qmd
Update code

```{r}
## ----- Libraries -------------------------------------------------------------
library(tidyverse)
library(bruneimap)
library(gt)
library(gtsummary)
library(lubridate)
library(GGally)

## ----- Load data sets --------------------------------------------------------
hsp <- 
  read_csv("data/hspbn_2024-12-12.csv") |>
  mutate(
    type = factor(type, levels = c("Detached", "Semi-Detached", "Terrace",
                                   "Apartment", "Land")),
    tenure = factor(tenure, levels = c("Freehold", "Leasehold", "Strata")),
    status = factor(status, levels = c("Proposed", "New", "Resale")),
    date = as.Date(date, format = "%d/%m/%y"),
    quarter = zoo::as.yearqtr(quarter)
  )
rppi <- 
  read_csv("data/rppi.csv") |>
  mutate(
    quarter = zoo::as.yearqtr(quarter),
    rppi = rppi / 100
  )
```

## Glimpse

```{r}
glimpse(hsp)
```

@@

```{r}
#| label: tbl-summary
#| tbl-cap: Summary of housing data
hsp |>
  mutate(
    price = price / 1000,
    storeys = factor(
      ifelse(storeys >= 3, "3+", as.character(storeys)),
      levels = c("1", "2", "3+")
    )
  ) |>
  tbl_summary(
    include = c(price:baths),
    by = district,
    missing = "no", 
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous2() ~ c("{mean} ({sd})", Range = "{min} - {max}", "{median} ({p25}, {p75})")
    ),
    label = list(
      price = "Price (BND 1,000)",
      type = "Property type",
      plot_area = "Plot area (acres)",
      floor_area = "Floor area (sq. ft.)",
      storeys = "Number of storeys",
      status = "Development status",
      tenure = "Land tenure",    
      beds = "Number of bedrooms",
      baths = "Number of bathrooms"
    ),
    digits = list(
      baths = 1,
      beds = 1
    )
  ) |>
  add_overall() |>
  add_n() |>
  modify_header(label ~ "**Variable**")
  # as_gt() |>
  # tab_options(
  #   quarto.disable_processing = TRUE
  # )
```

@@

```{r}
#| label: fig-spatial
#| fig-cap: Spatial distribution of median property prices by mukim.
#| message: false
#| warning: false
hsp_mkm <-
  hsp |>
  summarise(
    price = median(price, na.rm = TRUE, trim = 0.05),
    .by = mukim
  ) |>
  left_join(x = bruneimap::mkm_sf)

ggplot(hsp_mkm) +
  geom_sf(aes(fill = price), col = "gray50", linewidth = 0.5) +
  scale_fill_viridis_c(
    option = "cividis",
    na.value = "transparent",
    labels = scales::dollar,
    trans = scales::pseudo_log_trans(sigma = 0.001),
    name = "Median\nprice"
  ) +
  ggrepel::geom_label_repel(
    data = drop_na(hsp_mkm, price) |> 
      mutate(mukim = gsub("Mukim ", "", mukim)),
    aes(label = mukim, geometry = geometry),
    size = 2.7,
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0,       # Allow longer segments
    segment.size = 0.3,           # Adjust the thickness of lines
    segment.curvature = 0.1,      # Add curvature if needed
    force = 5                     # Increase force to spread labels
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw()
```

@@

```{r}
#| label: fig-price-evolution
#| fig-cap: "Median smoothed prices per square foot by property type using a 24-month (8-quarter) rolling window."
#| message: false
#| out-width: 100%
#| fig-height: 5
#| fig-width: 9
hsp_noland <- 
  filter(hsp, type != "Land") |>
  drop_na(type)
hsp_all <- hsp_noland
hsp_all$type <- "Overall"
hsp_sld <- 
  bind_rows(hsp_noland, hsp_all) |> 
  arrange(date)

slider::slide_period_dfr(hsp_sld, hsp_sld$date, "month", \(df) {
    df |>
    drop_na(type) |>
      summarise(
        date = min(date), 
        price = median(price, na.rm = TRUE),
        plot_area = median(plot_area, na.rm = TRUE),
        floor_area = median(floor_area, na.rm = TRUE),
        .by = type
      )
  }, .before = 18, .after = 6) |>
  distinct(date, type, .keep_all = TRUE) |>
  mutate(
    price_per_sqft = price / floor_area,
    type = factor(type, levels = c("Detached", "Semi-Detached", "Terrace",
                                   "Apartment", "Overall")) 
  ) |>
  ggplot(aes(x = date, y = price_per_sqft, col = type)) +
  geom_line(aes(linewidth = type)) +
  scale_x_date(
    breaks = scales::breaks_width("1 year"), 
    labels = scales::label_date("%Y"),
    name = NULL
  ) +
  scale_y_continuous(
    labels = scales::dollar,
    name = "Price per square foot (BND)"
  ) +
  scale_colour_manual(values = c(RColorBrewer::brewer.pal(4, "Set1"), "black")) +
  scale_linewidth_manual(values = c(rep(0.6, 4), 1.2)) +
  labs(
    col = NULL,
    linewidth = NULL
    # caption = "Median smoothed prices using a 12-month rolling window."
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) 
```

@@

```{r}
#| label: fig-rppi
#| fig-cap: "Comparison of quarterly median price per square foot indices (Median PPSF) and the official Residential Property Price Index (RPPI) from Brunei Darussalam Central Bank (BDCB)."
#| message: false
#| out-width: 100%
#| fig-height: 5
#| fig-width: 9
hsp_rppi <- hsp_noland |>
  drop_na(floor_area)

slider::slide_period_dfr(hsp_rppi, hsp_rppi$date, "month", \(df) {
    df |>
    drop_na(type) |>
      summarise(
        quarter = min(quarter),
        price = median(price, na.rm = TRUE),
        plot_area = median(plot_area, na.rm = TRUE),
        floor_area = median(floor_area, na.rm = TRUE)
      )
  }, .before = 1, .after = 1) |> 
  summarise(across(price:floor_area, \(x) median(x, na.rm = TRUE)), .by = quarter) |>
  mutate(
    price_per_sqft = price / floor_area,
    index = price_per_sqft / price_per_sqft[quarter == "2015 Q1"],
  ) |> 
  right_join(rppi) |>  
  pivot_longer(c(index, rppi), names_to = "series", values_to = "value") |>
  # mutate(series = factor(series, labels = c("PPS", "RPPI (BDCB)"))) |>
  ggplot(aes(x = quarter, y = value, col = series)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_line(linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent, name = "Index") +
  zoo::scale_x_yearqtr(
    format = "%Y-Q%q", name = NULL, #expand = c(0, 0),
    breaks = seq(2015, 2024.5, by = 0.25)
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  geom_text(
    data = tibble(
      quarter = c(2024.3, 2024.3),
      series = c("index", "rppi"),
      value = c(0.941, 0.955),
      label = c("Median PPSF", "RPPI (BDCB)")
    ),
    aes(label = label),
    hjust = 0
  ) +
  # directlabels::geom_dl(aes(label = series), method = "last.points") +
  coord_cartesian(xlim = c(2015, 2025))
```

>>>>

## Load data

```{r}
#| echo: true
#| eval: true 
#| file: data.R
```

@@

```{r}
#| label: tbl-summary
#| tbl-cap: Summary of housing data
hsp |>
  mutate(
    price = price / 1000,
    storeys = factor(
      ifelse(storeys >= 3, "3+", as.character(storeys)),
      levels = c("1", "2", "3+")
    )
  ) |>
  tbl_summary(
    include = c(price:baths),
    by = district,
    missing = "no", 
    type = all_continuous() ~ "continuous2",
    statistic = list(
      all_continuous2() ~ c("{mean} ({sd})", Range = "{min} - {max}", "{median} ({p25}, {p75})")
    ),
    label = list(
      price = "Price (BND 1,000)",
      type = "Property type",
      plot_area = "Plot area (acres)",
      floor_area = "Floor area (sq. ft.)",
      storeys = "Number of storeys",
      status = "Development status",
      tenure = "Land tenure",    
      beds = "Number of bedrooms",
      baths = "Number of bathrooms"
    ),
    digits = list(
      baths = 1,
      beds = 1
    )
  ) |>
  add_overall() |>
  add_n() |>
  modify_header(label ~ "**Variable**")
```

@@

```{r}
#| label: fig-spatial
#| fig-cap: Spatial distribution of median property prices by mukim.
#| message: false
#| warning: false
ggplot(hsp_mkm) +
  geom_sf(aes(fill = price), col = "gray50", linewidth = 0.5) +
  scale_fill_viridis_c(
    option = "cividis",
    na.value = "transparent",
    labels = scales::dollar,
    trans = scales::pseudo_log_trans(sigma = 0.001),
    name = "Median\nprice"
  ) +
  ggrepel::geom_label_repel(
    data = drop_na(hsp_mkm, price) |> 
      mutate(mukim = gsub("Mukim ", "", mukim)),
    aes(label = mukim, geometry = geometry),
    size = 2.7,
    stat = "sf_coordinates",
    max.overlaps = Inf,
    min.segment.length = 0,       
    segment.size = 0.3,           
    segment.curvature = 0.1,      
    force = 5                     
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw()
```

@@

```{r}
#| label: fig-price-evolution
#| fig-cap: "Median smoothed prices per square foot by property type using a 24-month (8-quarter) rolling window."
#| message: false
#| out-width: 100%
#| fig-height: 5
#| fig-width: 9
slider::slide_period_dfr(hsp_all, hsp_all$date, "month", \(df) {
  df |>
    filter(type != "Land") |>
    summarise(
      date = min(date), 
      price = median(price, na.rm = TRUE),
      plot_area = median(plot_area, na.rm = TRUE),
      floor_area = median(floor_area, na.rm = TRUE),
      .by = type
    )
}, .before = 18, .after = 6) |>
  distinct(date, type, .keep_all = TRUE) |> 
  mutate(
    price_per_sqft = price / floor_area,
    type = factor(type, levels = c("Detached", "Semi-Detached", "Terrace",
                                   "Apartment", "Overall")) 
  ) |>
  ggplot(aes(x = date, y = price_per_sqft, col = type)) +
  geom_line(aes(linewidth = type)) +
  scale_x_date(
    breaks = scales::breaks_width("1 year"), 
    labels = scales::label_date("%Y"),
    name = NULL
  ) +
  scale_y_continuous(
    labels = scales::dollar,
    name = "Price per square foot (BND)"
  ) +
  scale_colour_manual(values = c(RColorBrewer::brewer.pal(4, "Set1"), "black")) +
  scale_linewidth_manual(values = c(rep(0.6, 4), 1.2)) +
  labs(
    col = NULL,
    linewidth = NULL
    # caption = "Median smoothed prices using a 12-month rolling window."
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) 
```

@@

```{r}
#| label: fig-rppi
#| fig-cap: "Comparison of quarterly median price per square foot indices (Median PPSF) and the official Residential Property Price Index (RPPI) from Brunei Darussalam Central Bank (BDCB)."
#| message: false
#| out-width: 100%
#| fig-height: 5
#| fig-width: 9

hsp_rppi |>  
  pivot_longer(c(index, rppi), names_to = "series", values_to = "value") |>
  ggplot(aes(x = quarter, y = value, col = series)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_line(linewidth = 0.8) +
  scale_colour_brewer(palette = "Set1") +
  scale_y_continuous(labels = scales::percent, name = "Index") +
  zoo::scale_x_yearqtr(
    format = "%Y-Q%q", 
    name = NULL, 
    breaks = seq(2015, 2024.5, by = 0.25)
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  geom_text(
    data = tibble(
      quarter = c(2024.3, 2024.3),
      series = c("index", "rppi"),
      value = c(0.941, 0.955),
      label = c("Median PPSF", "RPPI (BDCB)")
    ),
    aes(label = label),
    hjust = 0
  ) +
  coord_cartesian(xlim = c(2015, 2025))
```

### 2024-12-14 manuscript.qmd
Update code

---
title: House price data in Brunei
authors:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: "https://haziqj.ml"
    corresponding: true
    affiliations:
      - name: Universiti Brunei Darussalam
        id: ubd
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410
  - name: Amira Barizah Noorosmawie
    affiliations:
      - ref: ubd        
  - name: Hafeezul Waezz Rabu
    affiliations:
      - ref: ubd
  - name: Lutfi Abdul Razak
    orcid: 0000-0003-1681-0895
    email: lutfi.razak@ubd.edu.bn
    affiliations:
      - ref: ubd              
date-modified: last-modified
nocite: |
  @*
abstract: |
  Abstract here.
---

@@

:::{.content-visible when-format="html"}

::: {.column-page }
{{< embed analysis.qmd#tbl-summary >}}
:::
:::

:::{.content-visible unless-format="html"}
{{< embed analysis.qmd#tbl-summary >}}
:::

@@

:::{.content-visible when-format="html"}
::: {.column-page }
{{< embed analysis.qmd#tbl-avail >}}
:::
:::

>>>>

---
title: House price data in Brunei
authors:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: "https://haziqj.ml"
    corresponding: true
    affiliations:
      - name: Universiti Brunei Darussalam
        id: ubd
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410
  - name: Amira Barizah Noorosmawie
    affiliations:
      - ref: ubd        
  - name: Hafeezul Waezz Rabu
    affiliations:
      - ref: ubd
  - name: Lutfi Abdul Razak
    orcid: 0000-0003-1681-0895
    email: lutfi.razak@ubd.edu.bn
    affiliations:
      - ref: ubd              
date-modified: last-modified
nocite: |
  @*
abstract: |
  Abstract here.
engine: knitr
---

```{r}
#| echo: false
#| eval: true 
#| file: data.R
```

@@

::: {.content-visible when-format="html"}
::: {.column-page}
{{< embed analysis.qmd#tbl-summary >}}
:::
:::

::: {.content-visible unless-format="html"}
{{< embed analysis.qmd#tbl-summary >}}
:::

@@

::: {.content-visible when-format="html"}
::: {.column-page}
{{< embed analysis.qmd#tbl-avail >}}
:::
:::
